---
title: "Conceptual and prosodic cues in child-directed speech can help children learn the meaning of disjunction"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Masoud Jasbi} \\ \texttt{masoudj@stanford.edu} \\ Department of Linguistics \\ Stanford University
    \And {\large \bf Akshay Jaggi} \\ \texttt{ajaggi@stanford.edu} \\ Department of Linguistics \\ Stanford University
    \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University
    }

abstract: 
    "Current theories of word learning focus on cues and mechanisms that help children's acquisition of open-class words such as nouns and verbs. Little is understood about the role of input in children's acquisition of closed-class words such as articles and connectives. This paper investigates the cues in child-directed speech that can help the interpretation and acquisition of the connective disjunction-word *or*. Study 1 uses a large corpus of parent-child interactions to show that despite low occurance of *or* in child-directed speech, children quickly reach the adult level of production for this word by age 4. Study 2 uses annotations on a subset of parent-child interactions to show that disjunctions in child-directed speech are accompanied by reliable cues to the correct interpretation (exclusive vs. inclusive). We present a computational model that learns from a handful of annotated examples to correctly predict the interpretation of a disjuntion. These studies suggest that reliable cues in child-directed speech can assist rapid acquisition of functional categories such as disjunction words."

keywords:
    "language acquisition; word learning; logical words; conjunction, disjunction."
    
output: cogsci2016::cogsci_paper
---

```{r global_options, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
library(knitr)
library(childesr)
library(tidyverse)
library(ggthemes)
library(lubridate)
library(magrittr)
library(lme4)
library(lmerTest)
library(forcats)
library(bootstrap)
library(jpeg)
```

<!-- 


Words with abstract meanings and multiple interpretations pose a challenge to children's acquisition of word meaning. Consider the word *or*. A disjunction such as ``*A or B*'' can be interpreted as inclusive (A or B, or both) or exclusive (A or B, not both). Linguistic investigations suggest that the core meaning of *or* is inclusive and that exclusive interpretations are the result of strengthening the inclusive meaning via (extra-semantic) factors such as pragmatic reasoning. However, research in language acquisition shows that while the majority of *or* examples children hear have an exclusive interpretation, in comprehension tasks children as young three understand the core meaning of *or* as inclusive disjunction. How can children quickly learn what they rarely hear? We conduct a large-scale quantitative study using an online corpus of child language and a small-scale annotation study to reveal the regularities in child directed speech that assist the interpretation of a disjunction. We find that exclusive interpretations of a disjunction are systematically correlated with conceptual and prosodic cues. *A or B* is interpreted as exclusive either when A and B are conceptually inconsistent, or when they receive a rise-fall intonation. We present a computational model that correctly predicts the exclusivity of a disjuntion given these cues from only a handful of examples. This study takes the inital steps in discovering the mechanisms that help children's acquisition of abstract function words.


Theories of word meaning mapping often consider forms that are mapped into concepts. Here we show a more complex picture where the form mapped must include prosody and mapping to cocnepts must take into account the conceptual consistency of the message for accurate learning from the input. The implication for usage-based approaches of word learning is that the picture is more complicated. The implication for nativist approaches is that input cues such as prosody and conceptual consistency carry a lot of information and may obviate the need for some innate principles and parameters.

How to say some important theoretical issues:
here we assume that the learner has the space of binary logical meanings available for mapping. Whether these concepts are innate as nativists assume or developed from other primitives is an issue we do not engage with. The main contribution of this paper is that once the space of logical connectives is available, it is not necessary to exclude exclusive disjunction as nativist accounts propose since that is learnable from the input.

--->

# Introduction

<!--
how do children learn the meaning of abstract function words? 
  We consider the case of "or"
  Our study suggests that mapping the meaning of "or" includes monitoring the formal and conceptual environment that it appears in
  
Acquisition facts: 
  children hear a lot more "and" than "or"
  children produce a lot more "and" and reach the adult rate of production very quickly for "and" but not for "or". they do not reach the adult rate of production for "or" at age 5 or 6. Suggesting that children learn the meaning/usage of "or" very gradually and that the logical inclusive meaning is late in acquisition
  children mostly hear the exclusive use of "or" from parents
  children's naturalistic productions show that they use it most often in its exclusive sense, mirroring what they hear most
  
Logical nativism: learning the meaning of disjunction is rapid and from early on adult-like
  humans have innate logical concepts for mapping to connective words
  this logical inventory for mapping logical words does not include exclusive disjunction
  children rely on this innate knowledge for mapping "or" rather than the structure of the input
  therefore, frequency does not predict acquisition
  Therefore, despite over-presence of the exclusive uses, children do not understand the meaning of "or" as exclusive
  comprehension tasks suggest children understand "or" as inclusive disjunction in a variety of environments

usage based approach:
  humans learn logical meanings from more primitive social concepts such as choice
  children's linguistic knowledge is shaped by the language they hear. they form usage rules and generalize.
  there is an expectation for frequency effect: more frequent words are learned earlier. 

Here we suggest that children's acquisition of disjunction is rapid, but likely due to systematic structure of the input that cues disjunction interpretation.

Study 1 shows that children reach the adult level of production for "and" and "or" relatively quickly. Study 2 shows that this rapid acquisition can be due to systmatic indicators to the interpretation of a disjunction in the input. This obviates the need for positing an innate ban on mapping to exclusive meaning.

-->

When we think of words, we often think of nouns like *cat*, adjectives like *red*, or verbs like *run*. These classes of words - called **content words** - have thousands of members and constantly add new ones like *Google*, *crunk*, or *chillax*. Their meanings often refer to concrete and tangible entities or events in the world. This is not the case for **function words** like *and*, *or*, *of*, and *the*. Functional classes have few members and rarely accept new ones. Their meanings are highly abstract and provide the glue that holds content words together to form complete thoughts. These properties make function words challenging for theories of word learning. While there has been a lot of research on cues and mechanisms that assist children's acquisition of content words, very little is known about cues that help children learn the meaning of function words. In this study, we make the initial steps by looking at the disjunction word *or* in parents' and children's speech.

The word *or* has long been a lab rat in linguistics and language acquisition. This interest in *or* is mainly due to its close connection to logical disjunciton. In formal logics, there are two types of disjunction: inclusive and exclusive. An **inclusive disjunction** such as "A $\vee$ B" is true when either A, B, or both are true. An **exclusive disjunction** such as "A $\oplus$ B" is true only when A or B is true but not both. The linguistic connective *or* can receive an inclusive interpretation like "A $\vee$ B" or an exclusive one like "A $\oplus$ B". For example, a waitor may ask if you would like something to eat or drink, not excluding the possibility that you would like both. However, the waitor may later ask if you would like to see the dessert menu or have the cheque, suggesting that you should chose one or the other. 

A large body of research in formal semantics and pragmatics seeks to specify the meaning of *or* and factors that determine its interpreation as inclusive and exclusive. @grice1975logicconvo argued that the core meaning of *or* is inclusive disjunction. The exclusive interpretation is the result of modifying this inclusive meaning by factors external to *or* itself. He suggested that upon hearing "A or B", we may exclude the possibility of both A and B being true simply because we reason that the speaker could have used the connective *and* instead of *or*. Therefore, the exclusive interpretation is the result of this pragmatic reasoning on the speaker's choice of *or* instead of *and* and not part of the meaning of *or* itself. @geurts2006exclusive argued that in many cases, exclusive interpretations stem from the inconsistent meaning of the options. For example, "to be or not to be" is exclusive simply because one cannot both be and not be! In an experimental study, @pruitt2013interpretation showed that in questions, exclusive interpretations are the result of a rise-fall intonation. In short, while the meaning of *or* itself is inclusive, the situation where both options are true may be excluded by pragmatic reasoning, inconsistent options, or a rise-fall intonation.

```{r aorb, fig.env="figure", fig.pos="b", fig.width=1.5, fig.align = "center" , fig.cap="Inclusive disjunction (A$\\vee$B) is true in situations where A, B, or both AB are true. Exclusive disjunction (A$\\oplus$B) is true in situations where only A or only B is true."}
library(jpeg)
#AorB <- png::readPNG("figs/AorB.png")
AorB <- jpeg::readJPEG("figs/AorB.jpg")
grid::grid.raster(AorB)
```

How do children learn the meaning of *or* given all this complexity in interpreting a disjunction? @morris2008logically investigated instances of *and* and *or* in parents' and children's speech. He found that *and* occurs about 3 times every 100 utterances in the speech of parents while *or* occurs only once or twice every 400 utterances. More importantly, children reach the adult level of production for *and* relatively quickly and around the age 3 but their productions of *or* do not reach the adult level in the first five years. He also found that the majority of *or* examples children hear (75-80\%) and produce (\%90) have an exclusive interpretation. Based on these findings, he concluded that children's acquisition of *or* is slow and they understand its meaning as exclusive before they add the inclusive meaning to their repertoire.

<!--The social media company LADbible reported the following in a tweet: "James Bond producer says next 007 could be black **or** a woman". A twitter user named Robert responded sarcastically with: "if only women could be black!" What in the producer's speech gave the impression that the next 007 could not be both black **and** a woman? The word *or*. A disjunction like "A or B" is associated with two interpretations: **inclusive**, and **exclusive**. "A or B" is inclusive when it is interpreted as "A or B **or both**". This is probably what LADbible meant when reporting the James Bond producer. However, "A or B" can also be exclusive: "A or B, but **not both**". Robert's response shows that he had an exclusive interpration of *or*. What factors determine the interpretation of *or* and how do children learn its meaning given the ambiguity it can give rise to?-->

<!--A large body of research in linguistics and philosophy in the past 50 years has created a common consensus on the meaning of *or* (see @Aloni2016). Data on the interpretation of disjunction across different sentences, contexts, and even languages show that the core meaning of disjunction words such as *or* is **inclusive**. This is similar to the definition of disjunction in formal logics. The **exclusive** interpretation of *or* is the result of enhancing its inclusive semantics via other (extra-semantic) factors such as intonation [@pruitt2013interpretation], inconsistency of the options [@geurts2006exclusive], and pragmatic reasoning over the speaker's choice of *or* instead of *and* [@grice1975logicconvo]. Therefore, interpreting a disjunction is a complex process that needs to take into account the meaning of *or* as well as different structural and contextual factors that accompany it. How do children learn such a complex interpretive system?-->

On constructivist accounts of the acquisition of disjunction, children learn the meaning of *or* by paying attention to how parents use it in different contexts. They form usage-rules and expand their usage repertoire as they grow up. The prediction is that children's production of *or* is slow and gradual, mirroring what they hear from parents. @morris2008logically found that *and* is about 13 times more frequent than *or* in parents' speech to children.  As predicted by the constructivist account, @morris2008logically reported that children also learn to produce *and* much more quickly than *or*. They reach the adult rate of *and* at age 3 while for *or* there is a gradual increase in production, possibly reaching the adult level at ages 5 or 6. The faster acquisition of *and* is consistent with the constructivist theory that emphasizes the role of usage frequency in children's linguistic development. @morris2008logically also reported that the majority of *or* examples children hear are exclusive. He reported that the majority of *or*'s children produce are also exclusive. Therefore, the inclusive semantics of *or* is developed after the exclusive interpretation.

However, several comprehension studies of *or* in different linguistic contexts show that children as young as three-years old interpret *or* as inclusive disjunction [@crain2012emergence]. This finding is surprising given @morris2008logically's finding that the majority of *or* examples children hear are exclusive. How do children learn the interpretation of disjunction as inclusive if they rarely hear it? We call this the **learning puzzle of disjunction**. @crain2012emergence suggests that instead of learning from the parents' usage of *or*, children rely on the innate knowledge that the disjunction operator in their native language must have an inclusive meaning. This nativist account predicts that *or* is learned relatively quickly and accurately by children.

Here we present an alternative answer that provides a synthesis between the nativist and constructivist accounts. We show that children can use regularities in parents' usage of *or* to learn the interpretation of disjunction from a handful of examples. In study 1, we use a large scale corpus study of parents and children's speech to show that both *and* and *or* appear relatively quickly in children's speech and reach the adult rate by the age 4. This finding is consistent with the comprehension studies that show children have an adult-like understanding of these words by the age four. In study 2, we replicate @morris2008logically's finding that the majority of *or* examples in child-directed speech have an exclusive interpretation. However, we also show that these exclusive interpretations correlate systematically with two factors external to *or*: intonation and consistency of the options. Exclusive interpretations are either inconsistent in nature (e.g. clean or dirty) or carry a distinct rise-fall intonation. We show that setting aside these cases, the interpretation of a disjunction is inclusive.

We argue that if children track the interpretive cues that accompany a disjunction, they can tease apart the role of the word *or* from factors that accompany and enhance it to shape the exclusive interpretation. This way children can discover that exclusivity corrleates with rise-fall intonation and inconsistent options, while *or* itself does not exclude the option of both disjuncts being true. We implement this novel account in a decision-tree that correctly learns to predict the interpretation of a disjunction with 80% accuracy after only a few examples. Our results show that the richness and systematicity of children's linguistic input allows rapid acquisition of disjunction with no need for an innate assumptions specific to the meaning of disjunction. We discuss the important implications of this work for the theories of word learning in the last section.



# Study 1: Corpus Study

```{r, childesdbImports, eval=FALSE, include=FALSE}
# This chunk uses the childesr package to import all English words from childes-db and create the processed files "raw_normalized" and "bySpeechAct_normalized"
# Since the file is too big, we only store two dataframes on github that are the results of running this chunk of code

# Import all words from childes-db
english_tokens <- get_tokens(collection = c("Eng-NA","Eng-UK"), 
                          corpus = NULL, 
                          role = c("target_child","Mother", "Father"),
                          age = NULL, 
                          sex = NULL, 
                          child = NULL,
                          token = "*")

#Convert and store children's age in years using the lubridate package
english_tokens$target_child_age_years <- 
  english_tokens$target_child_age %>% 
  duration("days") %>% 
  as.numeric("years")

#Take out data for the age range below 1 and above 6 years, this is because there is not much data in that range
english_tokens %<>% filter(target_child_age_years < 6, target_child_age_years > 1)

# Prepare the speech_act categories for this study based on the utterance_types in childes-db
# Categories: declarative, impertaive, question, and other
english_tokens$speech_act <-
  recode(english_tokens$utterance_type, 
         `broken for coding`="other",
          `imperative_emphatic` = "imperative",
         interruption = "other",
         `interruption question` = "question",
         `missing CA terminator` = "other",
         `no break TCU continuation` = "other",
         `question exclamation` = "question",
         `quotation next line` = "other",
         `quotation precedes` = "other",
         `self interruption` = "other",
         `self interruption question` = "question",
         `trail off` = "other",
         `trail off question` = "question"
         )

#####################################################################
## Now we're going to process the English words we have in "english_tokens" in two ways

######## 1. Just count all the instances of "and" and "or" and normalize them by the total number of words a speaker has said at a particular age (of the child)

# Group the tokens (words) by speaker role and the age of the child, count how many words for each speaker at each age
rawOverall_stats <-
  english_tokens %>%
  group_by(speaker_role, target_child_age_years) %>%
  summarize(total_freq = n())

# Group connective tokens by speaker and age of the child. count how many connective for each speaker at each age
rawConnective_stats <-
  english_tokens %>%
  filter(gloss=="and" | gloss=="or") %>%
  group_by(speaker_role, target_child_age_years, gloss) %>%
  summarize(freq = n())

# Normalize the connective count by dividing the connective frequency by the total word frequency for each speaker at each age
rawNormalized_stats <-
  full_join(rawConnective_stats, rawOverall_stats, by=c("speaker_role", "target_child_age_years")) %>%
  mutate(relFreq = freq / total_freq, relFreq_ppt = relFreq * 1000)

########## 2. Count all the instaces of "and" and "or" in DIFFERENT SPEECH ACTS and normalize them by THE TOTAL NUMBER OF WORDS IN SPEECH ACTS for each speaker at a particular age

#group the tokesn by speaker, age, and speech act, then count them
overall_stats <-
  english_tokens %>%
  group_by(speaker_role, speech_act, target_child_age_years) %>%
  summarize(total_freq = n())

# only take out "and" and "or"s, group them by speaker, age of the child, and speech act, count them
connective_stats <-
  english_tokens %>%
  filter(gloss=="and" | gloss=="or") %>%
  group_by(speech_act, speaker_role, target_child_age_years, gloss) %>%
  summarize(freq = n())

# now join the previous data frames together, divide the number of tokens in each group by the total number of words in that group
normalized_stats <-
  full_join(connective_stats, overall_stats, by=c("speech_act", "speaker_role", "target_child_age_years")) %>%
  mutate(relFreq = freq / total_freq, relFreq_ppt = relFreq * 1000)

########## Store both dataframes to be called later
write.csv(rawNormalized_stats, "2_processed_data/raw_normalized.csv")
write.csv(normalized_stats, "2_processed_data/bySpeechAct_normalized.csv")
```

```{r importProcessedData}
rawNormalized <- read.csv("2_processed_data/raw_normalized.csv")
bySpeechAct_normalized <- read.csv("2_processed_data/bySpeechAct_normalized.csv")
```

First, we conducted an exploratory investigation of *and* and *or* productions in parents and children. The goal of the study was to find out when children start producing these words and when they reach the adult rate of production. We conclude that children start producing *and* around 1.5 or 2 years of age, and *or* between the ages of 2 and 3. They reach the adult rate of production for *and* around 3 and for *or* around 4 or possibly earlier.

## Methods

We accessed the Child Language Data Exchange System ([CHILDES](https://childes.talkbank.org/), @macwhinney2000childes) via the online platform [childes-db](http://childes-db.stanford.edu/) and its associated R package childesr [@childesdb]. We extracted all instances of *and* and *or* from the English corpora (ENG-NA and ENG-UK). We limited our analysis to the data between one and six years because there is limited data outside this age range. We computed the relative frequency of connective production by dividing the total number of *and*/*or* in the speech of fathers, mothers, and children at a particular age by the total number of words spoken at that age. We present the relative frequency as parts per thousand.

```{r OverallConnectivePlots, fig.env="figure", fig.asp=0.7, fig.width=3.3, fig.show="hold", fig.pos="t", fig.cap= "The relative frequency of AND (top) and OR (bottom) per thousand words in the speech of fathers, mothers, and children between the ages of 1 and 6."}
rawNormalized %>%
  filter(gloss=="and") %>%
  ggplot(aes(target_child_age_years, relFreq_ppt, shape = speaker_role, color=speaker_role)) +
  geom_point(aes(), size=0.6) +
#  facet_grid(gloss, scales="free_y") +
  scale_y_continuous(limits = c(0, 75)) +
  scale_x_continuous(breaks=seq(1,6)) +
  labs(x = "", y="ANDs Per thousand words") +
  geom_smooth(aes(group = speaker_role, color=speaker_role), span=0.9) +
  scale_color_manual(values = c("seagreen3","seagreen", "darkblue")) +
  theme_few() +
  theme(legend.position=c(0.9, 0.8), text = element_text(size=9)) + 
  guides(shape=guide_legend(title=NULL), color=guide_legend(title=NULL))

rawNormalized %>%
  filter(gloss=="or") %>%
  ggplot(aes(target_child_age_years, relFreq_ppt, shape = speaker_role, color=speaker_role)) +
  geom_point(aes(), size=0.6) +
#  facet_grid(speech_act~gloss, scales="free_y") +
  scale_y_continuous(limits = c(0, 8)) +
  scale_x_continuous(breaks=seq(1,6)) +
  labs(x = "Child's Age (years)", y="ORs per thousand words") +
  geom_smooth(aes(group = speaker_role, color=speaker_role), span=0.9) +
  scale_color_manual(values = c("seagreen3","seagreen", "darkblue")) + 
  theme_few() +
  theme(legend.position=c(0.9, 0.8), text = element_text(size=9))+ 
  guides(shape=FALSE, color=FALSE)
```

## Results

```{r byspeechActPlots, fig.env="figure*", fig.asp=1.2, fig.width=3.3, fig.show="hold", fig.pos="t", set.cap.width=T, num.cols.cap=2, fig.align="center", fig.cap= "The relative frequency of AND (left) and OR (right) per thousand words in declaratives and questions in the speech of fathers, mothers, and children between the ages of 1 and 6."}
bySpeechAct_normalized %>%
  filter(speech_act == "declarative" | speech_act == "question", gloss=="and") %>%
  ggplot(aes(target_child_age_years, relFreq_ppt, shape=speaker_role, color = speaker_role)) +
  geom_point(aes(color=speaker_role), size=0.6) +
  facet_grid(speech_act~gloss, scales="free_y") +
  scale_y_continuous(limits = c(0, 60)) +
  scale_x_continuous(breaks=seq(1,6)) +
  labs(x = "Age (years)", y="Per thousand words") +
  geom_smooth(aes(group = speaker_role, color=speaker_role), span=0.9) +
  scale_color_manual(values = c("seagreen3","seagreen", "darkblue")) + 
  theme_few() +
  theme(text = element_text(size=9), strip.text.y = element_blank()) + 
  guides(shape=FALSE, color=FALSE)

bySpeechAct_normalized %>%
  filter(gloss=="or", speech_act == "declarative" | speech_act == "question") %>%
  ggplot(aes(target_child_age_years, relFreq_ppt, shape=speaker_role, color = speaker_role)) +
  geom_point(aes(color=speaker_role), size=0.6) +
  facet_grid(speech_act~gloss, scales="free_y") +
  scale_y_continuous(limits = c(0,7)) +
  scale_x_continuous(breaks=seq(1,6)) +
  labs(x = "Age (years)", y="") +
  geom_smooth(aes(group = speaker_role, color=speaker_role), span=0.9) +
  scale_color_manual(values = c("seagreen3","seagreen", "darkblue")) + 
  theme_few() +
  theme(text = element_text(size=9)) + 
  guides(shape=FALSE, color=FALSE)
```

In figure 1, we show the relative frequencies of *and* and *or* in the speech of parents and children between one and six years. In the speech of parents, *and* is produced around 20 times per thousand words while *or* is only produced around 2 times per thousand words. This confirms previous findings that *or* is much less frequent in child directed speech than a similar function word such as *and*.

*And* and *or* seem to show different developmental trajectories in the speech of children. For *and*, there is a rapid increase in its production between the ages of 1.5 and 3 before it reaches the adult rate around the age 3 and stay at that level until the age 6. For *or*, on the other hand, we see a slow incrase from the age 2 until at least age 6, as it approaches the adult rate. This difference in the development of *and* & *or* production was attributed to the frequency of these items in child-directed speech. Since *and* is much more frequent than *or*, it is learned much faster than *or*. @morris2008logically argued that such patterns support the item-based and usage-based acquisition of logical words.

However, the analysis above does not control for other factors that can affect the production of words by children. An important factor to control for is the development of speech acts. While content words such as *dog* or *chair* may appear freely in different types of speech acts, function words are highly constrained by the type of speech acts they can appear in. For example, it is reasonable to assume that question words such as *how* and *why* are much more likey to occur in questions than statements (declaratives). If parent-child interaction is such that parents ask more questions than children, it is not surprising to find higher rates of *how* and *why* production in parents than children. Therefore, it is important to control for the speech act a function word appears in.

Figure 2 shows the relative frequencies of *and* and *or* in questions vs. declaratives, in the speech of parents and children between one and six years. Here, the relative frequency is computed by dividing the total number of *and*/*or* in a question/declarative in the speech of fathers, mothers, and children at a particular age, by the total number of words in a question/declarative spoken at that age. As before, we present the relative frequency as parts per thousand.

The results show similar developmental trajectories for the production of *and* and *or* in children. For both words, there is a relatively rapid incease in their frequency between the ages of 2 and 4 before reaching the parent rate at the age of 4 and staying at that rate until the age of 6. This pattern of production is consistent with the nativist observations that the acquisition of *and* and *or* is rapid and that children have an adult-like comprehension of these two connectives at the age 4. 

## Summary

In this study we found that *and* is a lot more frequent than *or* in child directed speech. Looking at the relative frequency of *and* and *or* between the ages of 1 and 6, it appeared as if children show radically different developmental trajectories for these two connectives. Children's production of *and* seemed to sharply increase around the age 1.5 and reach the adult rate around age 3 while *or* production increased slowly and linearly until it reached the adult rate at age 6. However, we showed that this discrepency is to a large part due to the distribution of function words in different speech acts. Since *or* is more frequent in questions and children produce fewer questions than parents, the frequency of *or* in children's speech is underestimated if we normalize by the total number of words they produce. Instead, when we look at the production of *and* and *or* in specific speech acts - declaratives and questions - and normalize by the total number of words produces in each speech act, we see that *and* and *or* show similar developmental patterns. We argue that the developmental trajectory of *and* \& *or* is best described as a quick increase in production between the ages of 2 and 4 and staying around the parents' rate between the ages 4 and 6. This is compatible with the comprehension studies which suggest children understand *and* and *or* by the age four. 

```{r warmup, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggthemes)
library(bootstrap)
library(forcats)
library (irr)
library(caret)
library(lme4)
library(lmerTest)
library(knitr)
library(ggplot2)

## for bootstrapping 95% confidence intervals
theta <- function(x,xdata) {mean(xdata[x])}
ci.low <- function(x) {
  quantile(bootstrap(1:length(x),1000,theta,x)$thetastar,.025)}
ci.high <- function(x) {
  quantile(bootstrap(1:length(x),1000,theta,x)$thetastar,.975)}
```

```{r data_prep, echo=FALSE, warning=FALSE, message=FALSE}
alex_data <- read.csv("1_raw_data/ProvidenceData-alex.csv", nrows=100)
lily_data <- read.csv("1_raw_data/ProvidenceData-lily.csv", nrows=99)
vio_data <- read.csv("1_raw_data/ProvidenceData-vio.csv", nrows=102)
will_data <- read.csv("1_raw_data/ProvidenceData-will.csv", nrows=100)
naima_data <- read.csv("1_raw_data/ProvidenceOR-Naima.csv", nrows=241)

naima_data <- 
  naima_data %>%
  filter(!is.na(exclusivity), !is.na(intonation))

all_data <- rbind(alex_data, lily_data, vio_data, will_data, naima_data)

all_data$intonation <- as.factor(all_data$intonation)

all_data$exclusion <- fct_recode(all_data$exclusion, "Consistent" = "ELS", "Inconsistent" = "EXC")
all_data$intonation <- fct_recode(all_data$intonation, "Flat" = "0", "Rising" = "1", "Rise-Fall" = "2")
all_data$intonation <- fct_recode(all_data$intonation, "Flat" = "0", "Rising" = "1", "Rise-Fall" = "2")

```

# Study 2: Annotation Study

Study 1 showed that even though *or* is not very frequent in child-directed speech, children learn and produce it relatively quickly and reach the adult rate around the age 4. In Study 2, we conducted a detailed and small-scale investigation of *or* productions in child-directed speech. The goal of the study was to find out how children learn the meaning of *or* from such little data. We conclude that children hear significantly more exclusive uses of *or* supporting @morris2008logically's findings. However, we also find that the exclusive uses of *or* are rich in structures that index the exclusive interpretation. These indicative structures include rise-fall intonation and consistency of the options. Exclusive interpretations are either inconsistent in nature (e.g. clean or dirty) or carry a distinct rise-fall intonation. We then show that a decision tree model can learn to accurately (>80%) predict exclusivity using this information after seeing few (<20) examples. Instances of *or*  without these indices are most likely interpreted as inclusive. 

## Methods 

We accessed [the Providence corpus](https://phonbank.talkbank.org/browser/index.php?url=Eng-NA/Providence/) [@demuth2006word] via the phonbank section of [the TalkBank system](https://talkbank.org/). We extracted all instances of *or* along with the two utterances before and after to provide context using [the CLAN software](http://alpha.talkbank.org/clan/). We annotated the examples for three major categories: 1. Interpretation 2. Intonation and 3. Conceptual consistency. Table 1 shows these categories along with their subcategories and an example for each subcategory. 

### Disjunction 
The first category - disjunction interpretation - was our dependent measure. Annotators listened to a disjunction like "A or B" and decided whether the speaker intended to imply "not both A and B" (exlcusive) or "possibly both A and B" (inclusive).

### Intonation 
For the second category - intontation - annotators listened to the sentence containing disjunction and decided whether the intonation contour on the disjunction is rise-fall, rise, or flat. Table 1 includes examples that are prototypically read aloud with the intonation contour they are subcategorized as. 

### Consistency
Finally, for conceptual consistency, annotators decided wether the propositions that make up the disjunction are inconsistent. Our annotators used the following diagnostic to decide the consistency of the disjuncts: Two disjuncts were marked as inconsistent if replacing the word *or* with *and* produced a contradiction. For example, changing "the ball is in my room and your room" to "the ball is in my room and your room" produces a contradiction because the propositions cannot be both true at the same time. In such cases, the inclusive meaning is simply not available due to the nature of the propositions themselves and not the interpretation of the connective. It is important to note here that this is a particularly strict diagnostic. In many cases, the possibility of both propositions being true is ruled out based on prior knowledge and expectations of the situation. For example, when asking people whether they would like tea or coffee, it is often assumed that people choose one or the other. However, wanting to drink both tea and coffee is not conceptually inconsistent. It is just very unlikely. Our annotations of consistency are very conservative in that they consider such unlikely cases as consistent.

```{r annotatuin_table, results="asis"}
annotation_categories <- data.frame(
  Category = c("Interpretation", "", "Intonation", "", "", "Consistency", ""),
  Subcategory = c("Exclusive", "Inclusive", "Flat", "Rise", "Rise-Fall", "Consistent", "Inconsistent"),
  Examples = c("Wanna stay or go?", "Anything to eat or drink?", "I'll get tea or coffee.", "Anything to eat or drink?", "Wanna stay or go?", "I'll get tea or coffee.", "Wanna stay or go?")
)

annotation_table <- xtable::xtable(annotation_categories, 
                      caption = "Annotation categories and examples.")

print(annotation_table, type="latex", comment = F, table.placement = "H", include.rownames=FALSE, hline.after = c(0,2,5))
```

Finally, to test inter-rater reliability, the two raters annotated the same 240 utterances. The interrater reliability was calculated over 8 iterations of 30 examples each. Training only completed after 3 consecutive iterations had substantial agreement between the raters (Cohen's $\kappa > 0.7$) for all categories.

```{r data_wrangling, echo=FALSE, warning=FALSE, message=FALSE}
raw_data <- 
  all_data %>%
  group_by(id, exclusivity) %>%
  summarise(counts= n()) %>%
  spread(exclusivity, counts) %>%
  replace(is.na(.), 0) %>%
  mutate(total = IN + EX) %>%
  gather(EXIN, counts, EX:IN) %>%
  mutate(prop = counts / total) %>%
  group_by(EXIN) %>%
  summarize(cih = ci.high(prop),
            cil = ci.low(prop),
            prop = mean(prop))

graph_data <- 
  all_data %>%
  group_by(id, intonation, exclusion, exclusivity) %>%
  summarise(counts= n()) %>%
  spread(exclusivity, counts) %>%
  replace(is.na(.), 0) %>%
  mutate(total = EX + IN) %>%
  gather(exclusivity, counts, EX:IN) %>%
  mutate(prop = counts / total) %>%
  group_by(exclusivity, intonation, exclusion) %>%
  summarize(cih = ci.high(prop),
            cil = ci.low(prop),
            prop = mean(prop)) 

counts_table<-  
  all_data %>%
  group_by(intonation, exclusion, syn_level, exclusivity) %>%
  summarise(counts= n()) %>%
  spread(exclusivity, counts) %>%
  replace(is.na(.), 0) %>%
  mutate(total = IN + EX)

exclusivity_overall <-
  all_data %>%
  group_by(exclusivity) %>%
  summarise(counts= n()) %>%
  spread(exclusivity, counts) %>%
  replace(is.na(.), 0) %>%
  mutate(total = IN + EX) %>%
  gather(exclusivity, counts, EX:IN) %>%
  mutate(prop = counts / total)

cds_disjunction_model <- summary(glmer(exclusivity ~ intonation + exclusion  + syn_level + (1|id), family="binomial", data=all_data))
```


```{r interpretation, fig.env='figure',fig.width=2, fig.height=2, fig.show='hold', fig.pos="b", set.cap.width=T, num.cols.cap=2, fig.align="center",fig.cap = "Proportion of exclusive and inclusive interpretations of disjunction in child-directed speech. Error bars represent bootstrapped 95\\% confidence intervals"}
ggplot(raw_data, aes(x= EXIN, y=prop, fill=EXIN)) + 
  theme_few(base_size = 10) + 
  geom_col(width=0.7) +
  geom_linerange(aes(ymax = cih, ymin = cil)) + 
  guides(fill=FALSE) +
  labs(title = "", x = "", y = "") + 
  scale_colour_solarized()
```
 
## Results 

```{r interpretationByIntonationAndConsistency, fig.env='figure*',fig.pos='t',fig.width=4.2, fig.height=3, fig.align="center",fig.show='hold',set.cap.width=T, num.cols.cap=2, fig.cap = "Distribution of exclusive and inclusive interpretations broken down by intonation (flat, rise, rise-fall) and consistency. Error bars represent bootstrapped 95\\% confidence intervals"}
ggplot(graph_data, aes(x= exclusivity, y=prop, fill=exclusivity)) + 
  geom_col(width=0.7) +
  geom_linerange(aes(ymax = cih, ymin = cil)) +
  guides(fill=FALSE) +
  theme_few(base_size = 10) + 
  labs(title = "", x = "", y = "") + 
  scale_colour_solarized() + facet_grid(exclusion~intonation)
```

First, similar to Morris (2008), we found that the majority of *or* examples in CDS receive an exclusive interpretation ($\sim$%`r 100 * round(exclusivity_overall$prop[1],2)`). Figure 3 shows this difference in distribution. However, the rate of exclusive interpretations change systematically when we break the data down by prosody and consistency (Figure 4). Given a rise-fall intonation contour, a disjunction is almost always interpreted as exclusive. Similarly, if the propositions are inconsistent, the disjunction is most likely interpreted as excluisive. When either of these two features are absent, a disjunction is more likely to receive an inclusive interpretation.

A mixed-effects binomial logistic regression using the pakcage \{lme4\} [@bates2014lme4] in R with the fixed effects of intonation and consistency, as well as the random effects for children, found both intonation and consistency significant in interpreting disjunctions. Disjunctions were more likely to be interpreted as exclusive if they received a rise-fall intonation ($\beta$=`r round(cds_disjunction_model$coefficients[3],2)`, $z$=`r round(cds_disjunction_model$coefficients[11],2)`, $p < 0.001$) or if they were inconsistent ($\beta$=`r round(cds_disjunction_model$coefficients[4],2)`, $z$=`r round(cds_disjunction_model$coefficients[12],2)`, $p < 0.001$). Disjunctions were more likely to be interpreted as inclusive if they were consistent and received a rising ($\beta$=`r round(cds_disjunction_model$coefficients[2],2)`, $z$=`r round(cds_disjunction_model$coefficients[10],2)`, $p < 0.001$) or flat intonation ($\beta$=`r round(cds_disjunction_model$coefficients[1],2)`, $z$=`r round(cds_disjunction_model$coefficients[9],2)`, $p < 0.001$).

```{r treeDiagram, fig.env = "figure", fig.pos = 'H', fig.align='center', fig.width=3, fig.height=2,fig.cap="Optimal decision tree training on 100 datapoints. Provides series of two binary decisions to decide exlcusivity interpretation. Intonation $>$ 1.5 are rise-fall while intonation $<$ 1.5 are flat or rising. Consistency $>$ 0.5 are consistent while consistency $<$ 0.5 are  inconsistent."}

treeDiagram <- png::readPNG('figs/treeDiagram.png')
grid::grid.raster(treeDiagram)
```

Next, using Sci-kit Learn's Decision Tree Module [@pedregosa2011scikit], we built a predictive model to train on annotated *or* utterances and predict the exclusivity of unseen *or* utterances (annotated for intonation and consistency). We randomly sampled 100 examples for training and 300 examples for testing. Averaged over 100 trials, the average accuracy of a binary tree was 83%. More remarkably, the tree achieved an average of 80% accuracy after training on only 20 examples. The control flow of the average decision tree on a single example is as follows. If *or* has neither rise-fall nor inconsistent disjuncts, it is marked inclusive. Otherwise, exclusive. The success of such a simple tree indicates that children could use a simple model to rapidly learn the exclusive interpretation of *or* from little data. 

```{r learningCurve, fig.env = "figure", fig.pos = 'H', fig.align='center', fig.width=3, fig.height=3,fig.cap="Decision Tree Accuracy as a function of number of examples seen. Tested on a constant 300 examples. Dashed line marks 80% accuracy threshold."}
learning_curve <- png::readPNG('figs/learning_curve.png')
grid::grid.raster(learning_curve)
```

## Summary 

In study 2, we confirm @morris2008logically's finding that exclusive interpretations of *or* are far more common than inclusive interpretations. However, we also show that the majority of these exclusive interpretations coincide with systematic indicators. Disjunctions that are accompanied by rise-fall intonation or inconsistent disjuncts are far more likely to be exclusive. Disjunctions that do no bear these features are more likely to be inclusive. Accounting for these external factors, a simple decision tree can rapidly learn to predict the exclusive interpretation of the disjunction. 

# Discussion

Studies presented in this paper resolve two puzzles in children's acquisition of disjunction. First, previous studies suggested a discrepency between children's production of *or* and their comprehension. Comprehension studies showed that children have an almost-adult-like interpretation of *and* and *or* at around the age 4. However, production studies suggested that *and* is learned and produced quickly while children are slower in reaching the adult rate of production for *or*. Study 1 showed that when we control for the environments (speech acts) that these connectives are suitable for, we observe that both *and* and *or* are learned relatively quickly and produced at an adult rate around the age 4. This is compatible with the comprehension studies that show 3-to-5-year-olds understand the meaning of *or* as inclusive disjunction and *and* as conjunction. 

The second puzzle that this study addressed was the (apparent) discrepency between what they hear from parents and the knowledge they manifest in comprehension studies. Previous studies showed that the majority of *or* examples children hear receive an exclusive interpretation, yet in comprehension tasks they interpret *or* as inclusive disjunction. Study 2 showed that even though the majority of *or* examples are exclusive, this exclusivity is due to prosodic and conceptual properties of the disjunction and not *or* itself. We showed when we break down interpretations of disjunction in child-directed speech by prosody and concpetual consistency, the vast majority of exclusive interpretations are either inconsistent conceptually (e.g. clean or dirty) or are accompanied by a distinctive rise-fall intonation. Otherwise, a disjunction that does not bear these properties is more likely to be interpreted as inclusive. This finding suggests that if children break down disjunction interpretations by cues that accompany them, they can learn the adult like interpretation of disjunction from child-directed speech. We showed that a decision tree can use prosody and conceptual consistency to identify the correct interpretation using less than 20 examples.

The account presented in this paper has implications for the theories of form-meaning mapping in general and more specifically theories of function word acquisition. Form-meaning mapping in child language acquisition is often construed as the task of associating a novel and isolated form such as *gavagai* to a delimited concept such as rabbit and storing it in memory. However, the case of disjunction paints a more complicated picture. For disjunction, it is not enough to isolate the word *or* for mapping and the learner needs to also consider the prosodic contour that accompanies *or*. When mapping to meaning, it is not enough to simply consider the overall interpretation and the learner should take into account the conceptual properties of the accompanying words and phrases. In other words, mapping the form *or* to its meaning involves considering both other forms as well as concepts that accompany *or* and its core meaning.

Finally, the acquisiton of *or* is important for form-meaning mapping in function words such as determiners (e.g. *the*, *a*), connectives (e.g. *and*, *or*, *if*), and auxiliaries (e.g. *might*, *should*, *can*). The word learning literature has mostly focused on the acquisition of content words such as nouns, adjectives, and verbs.  Theories of word learning for content words often rely on the physical and observable cues present in the child's surroundigs to learn the relevant meanings. Such theories face difficulty when dealing with function words since most function words do not have observable cues to their meanings. The research presented here provides early indication on the types of cues that can help children learn the meaning of function words.

<!--Finally, we would like to add that conceptual consistency and rise-fall prosody are not the only cues that help disjunction interpretation. The propositions in the LADbible tweet discussed at the beginning of the paper do not carry a rise-fall intonation and are not inconsistent either. Instead the pragmatic reasoning on why the word *or* was used and what is being discussed gives rise to an exclusive interpretation.--> 

# Acknowledgements

We would like to thank Eve Clark for her comments and guidance with this project. We would also like to thank Kutay Serova and Salma Sebt for their help.

# References 

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent
